---
title: "Stroop Test"
author: "Tarique Hasheem"
date: "12/20/2017"
output: html_document
---

#### <font color="#00ccff">Background</font>

<font color="grey">In psychology the Stroop effect is an interesting occurrence 
that demonstrate reaction times of participants who are asked to say the color 
of a word that's presented to them out loud.  For example <font color="red">red
</font>, <font color="green">green</font> known as *congruent words* where the 
color and words match. The opposite are *incongruent words* where the words are 
not in the color they appear, for example <font color='orange'>blue</font> or 
<font color='purple'>grey</font>.  The results are fascinating, reaction times 
to incongruent words are slower than congruent words.  In our project we'll 
investigate the Stroop effect using statistics to help uncover some of our 
questions.

#### <font color="#00ccff">Questions for Investigation</font>

#### 1.  What is our independent variable?  What is our dependent variables?

Our independent variable are the colors and our dependent variable are 
reaction time.  

#### 2.A.  What is an appropriate set of hypotheses?  

μ = population mean

There is no difference in population means of response time under incongruent 
and congruent scenarios, which can be mathematically represented as 
H0: μC = μI.

There is a difference in population means of response time under incongruent and
congruent scenarios, which can be mathematically represented as
Ha: μC ≠ μI.

##### 2.B.  What kind of statistical test do you expect to perform?

The dependent t-tests compares the mean of two groups to determine whether 
there is a statistically significant difference between them.  Dependent 
test indicates that the same participant have been tested more than once.  
T-test's are great when we don't know the population parameters.  We're 
interested in a two-tailed test to simply test for the possibility that of 
the relationship in both directions.  In other words we want to know if 
reaction times are decreased or increased or no significant change in reaction 
time. 

#### 3.  Below are measures of centrality.

```{r}

df <- read.csv("stroopdata.csv")
conMean <- round(mean(df$Congruent), 2)
print(paste("Congruent Maean:", conMean))
```


```{r}
inconMean <- round(mean(df$Incongruent), 2)
print(paste("Incongruent Mean:", inconMean))
```


Below are the variance.

```{r}
conVar <- round(var(df$Congruent), 2)
print(paste("Congruent Variance", conVar))
```


```{r}
incVar <- round(var(df$Incongruent), 2)
print(paste("Incongruent Variance", incVar))
```

Below are the standard deviation for Congruent.

```{r}
con_sd <- round(sd(df$Congruent), 2)
print(paste("congruent standard deviation", con_sd))
```

Below is the standard deviation for In congruent.
```{r}
incon_sd <- round(sd(df$Incongruent), 2)
print(paste("incongruent standard deviation", incon_sd))
```

#### 4. Visualizations

Below we have tow box plots, one for our congruent data set and the second for 
our incongruent dataset.  We can clearly see the outliers and median scores for
both datasets.

```{r echo=FALSE, warning=FALSE}

library(ggplot2)
library(dplyr)

congruent_DataSet <- df$Congruent
incongruent_DataSet <- df$Incongruent

congruent_DataSet <- df$Congruent
incongruent_DataSet <- df$Incongruent
boxplot(congruent_DataSet, horizontal=TRUE)


```

```{r}
boxplot(incongruent_DataSet, horizontal=TRUE)
```



#### 5. Statistical Test

Dependent t-test

Lets print the mean for both our congruent and incongruent datasets.

```{r}
print(paste("congruent mean", conMean))
print(paste("incongruent mean", inconMean))
```

Our sample size is n = 24

What is our point estimate for congruent mean minus incongruent mean?

```{r}
point_estimate = round((conMean - inconMean),2 )
print(paste("Our point estimate is:", point_estimate))
```

Now we need the difference between the congruent and incongruent tests 
results.  We can simply create a new column in our dataset with the results.  
We'll print the results below.

```{r}
con <- congruent_DataSet
incon <- incongruent_DataSet
df["difference"] <- round((con - incon), 3)
diff = df$difference
print(paste(diff))
```

Now lets subtract the difference from our point estimate and square our results.
We'll print the results below.

```{r}
df["diff_minus_pointEstimate_squared"] <- ((diff - point_estimate)^2)
print(paste(round(df$diff_minus_pointEstimate_squared, 3)))
```

Now we will take the figures from the cell above (and our new column in
dataframe) and divide it by n-1.

```{r}
a = sum(df$diff_minus_pointEstimate_squared)/(24-1)
print(paste(round(a, 2)))
```

Let's now get the square root.

```{r}
b = sqrt(a)
print(paste(round(b, 2)))
```

Our sample standard deviation of the differences is 4.86.  Now lets calculate
the t-statistic.  

```{r}
s = (-7.97) / (4.86 / sqrt(24))
print(paste("s =", round(s, 2)))
```

Is there a significant difference between congruent and incongruent response
times at an alpha level of 0.05?  What are the t-critical values?

Our t-critical values are +/- 2.069.  Since our t-statistic score (-8.03) is 
greater than our t-critical score, we reject the null hypothesis.  In other
words, there is a significant difference in response time for congruent and 
incongruent words.

We will now calculate the effect size measure with Cohen's d.

```{r}
cd <- -7.97 / -8.03
print(paste("Cohens d", round(cd, 3)))
```

What is our confidence interval?

```{r}
t_critical <- 2.069
ci_min <- (point_estimate) - (t_critical*(s/sqrt(24)))
ci_max <- (point_estimate) + (t_critical*(s/sqrt(24)))
print(paste("CI", round(ci_min, 2),",", round(ci_max, 2)))
```

The CI interval is telling us that response times are increased by 4.58 to 11.36
longer with incongruent words.


</font>

***

<br></br>
<br></br>